Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-12 draft/ConFut.tex
now it 9 - appril and after club day
after grammer correct now will print to review

\chapter{Discussion and Conclusion}

\section{Conclusion}
\label{sec:ConclusionConclusion }

This paper presented a new approach to sketch recognition using PSO. It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system. The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}. The trade off between accuracy achieved and time complexity must be further investigated to achieve better results.  The system introduced an efficient method to sketch recognition using the PSO. The test performed confirmed that PSO achieve better performance and optimization than other algorithms. The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system. The only draw back of the system that it takes more time with comparison with similar algorithms.

@@

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}
The next step in this research is to complete the clustering algorithm for fully automated sketch recognition. The clustering must be performed without the user explicit involvement.  The currently used algorithm must be modified to accept error reporting from users. The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve. Other area of modification can be the features extraction and classifier. Adding more spatial features and structure features will improve classifications. After segmenting the stroke the system can generate a structural graph from segments and there spatial relations. This graph will be matched with the set of known symbol template to make the identification. Those result along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

>>>>

\chapter{Discussion and Conclusion}
\label{sec:DiscussionConclusion}

\section{Conclusion}
\label{sec:ConclusionConclusion}

This paper presented a new approach to sketch recognition using PSO. It was noted that the PSO in general improve the accuracy of the final symbol recognition in the system. The use of both speed and curvature data helps in improvement of the PSO algorithm over the original algorithms \cite{CruveDivisionSwarm,PolygonApproximationPSO}. The tradeoff between accuracy achieved and time complexity must be further investigated to achieve better results.  The system introduced an efficient method to sketch recognition using the PSO. The test performed confirmed that PSO achieve better performance and optimization than other algorithms. The number and the complexity of symbols recognized by the system varied from simple to complex but did not affect the final recognition of the system. One drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.  Another drawback is that until now we only applied the system to single symbols.  Full free hand sketch test has not been applied on the system. %sketched

@@

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}
The next step in this research is to complete the clustering algorithm for fully automated sketch recognition. The clustering must be performed without the user explicit involvement.  The currently used algorithm must be modified to accept error reporting from users. The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve. Other area of modification can be the features extraction and classifier. Adding more spatial features and structure features will improve classifications. After segmenting the stroke the system can generate a structural graph from segments and there spatial relations. This graph will be matched with the set of known symbol template to make the identification. Those results along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

### 2009-04-12 draft/Results.tex
after grammer correct now will print to review

\section{Datasets}
\label{sec:Datasets}
Two different data set were used, 
%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from three different domains ( UML, mechanical and electrical engineering symbols). Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style,size or orientations. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}
 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption{The Symbol Set}
	\label{fig:symbolSet}
\end{figure}

@@

Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}). The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error. Figure \ref{fig:PossibledpPreCalculation} shows the average number of Possible dominant point $P_{pd}$ with respect to information used in the system. The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmentation stage decreases the \textit{Segmentation Error}. It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone. This is because we found out that there was some redundant points found using more than one information.

@@

 The effect of number of agents, the error thresholds and various other parameters was investigated. The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and  \ref{fig:swarmtestingS2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increases the \textit{Segmentation Error} value decrease. Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1} . Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}). The final values are for maximum number of iteration = 80 with 15 particles. The final number of agents and maximum swarm iteration used in the system was based trade off between error achieved versus the computation time.

@@

%The fig. \ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration dose not affect the error calculated. Similar behavior is noticed in fig. \ref{fig:datapopulation.jpg}.  These curves and test result in choosing the swarm parameters $c_1,c_2$ and maximum number of iteration and the swarm population. The final values are for maximum number of iteration = 150 with 15 particle for better compensation in the time domain.

@@

\subsection{Segmentation Algorithms}
\label{sec:SegmentationAlgorithms}
The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}. These result shows the originals stroke and the segmentation that the system generated for the stroke. The figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.  Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes. It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmentation algorithm on the recognition accuracy.

@@

e performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse detection module appears to be superior. The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm. %show that combining algoirthmsachieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

@@

The result illustrated in this chapter shows the work done on the system and experiments performed. The results shows that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.

>>>>

\section{Datasets}
\label{sec:Datasets}
Two different data set were used, 
%For classifier and algorithm testing the system used a data set collected by
\begin{description}
	\item [Hse Dataset] (denoted as Hs-DB) Data set collected by \cite{HeloiseBeautification} and was previously used by \cite{Oltmans07} as benchmark for sketch recognition problem. The data are drawn by 16 users each of them have drawn from 30 to 50 samples for each shape. Figure \ref{fig:symbolSet} shows the shapes used in the data set. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. In worth noting that \cite{HeloiseBeautification} and \cite{Oltmans07} achieved 96.7\% and  94.4\%  on this dataset respectively. %, 94.4\%  , which
%is comparable to the 96.7%  %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols.
		\item [Multi Domain Dataset] (denoted as MD-DB) A dataset we collected manually from different users. The dataset contains symbols from three different domains ( UML, mechanical and electrical engineering symbols). Data are collected from 4 different users each of them have drawn from 20 to 40 samples of each shapes. Figures \ref{fig:LogicSet} and \ref{fig:ElectSet} show some of the shapes used in the dataset. Symbols from different domains were gathered to evaluate the system throughout different types of shapes. Users were requested to draw symbols without any restriction on style, size or orientations. The dataset was divided into training set and test set. Four different splits for the training and test data are generated from the dataset. %The results displayed are the average recognition accuracy of the four splits resulted from the classifier after recognizing the symbols. 
\end{description}
 %We divided the dataset equally into training set and test set.  The results displayed are the recognition accuracy resulted from the classifier after recognizing the symbols. \\
%We divided the dataset into four sets each set contains four different users. Furthermore, each set is divided into train and test set where the trained set where used to train the classifiers and the test set where used for testing.  The algorithms are then tested using the four sets and the results are averaged.
\begin{figure}[]\centering
\fbox{  \parbox{8cm}{% 
		\includegraphics[scale=0.7]{images/symbolSet.eps}	}
		}
	\caption{The Symbol Set}
	\label{fig:symbolSet}
\end{figure}

@@

Figure \ref{fig:SegErrPreCal} shows the \textit{Average Segmentation Error} using different preliminary calculation (Section \ref{sec:CurvatureCalculation}). The figure shows that using all computations (time difference, direction, speed and curvature) for each stroke decrease the Segmentation error. Figure \ref{fig:PossibledpPreCalculation} shows the average number of Possible dominant point $P_{pd}$ with respect to information used in the system. The results shows that the using all information increases the number of Possible dominate point $P_{pd}$ which in segmentation stage decreases the \textit{Segmentation Error}. It is worth noting that when using all information the number of Possible dominant point $P_{pd}$ is not the same as the sum of using every information alone. This is because we found out that there was some redundant points found using more than one information source.

@@

 The effect of number of agents, the error thresholds and various other parameters was investigated. The result of these tests are shown in Figure \ref{fig:swarmtestingS1} and \ref{fig:swarmtestingS2}. Figure \ref{fig:swarmtesting} shows that as the number of agents increase the \textit{Segmentation Error} value decrease. Similar results for the maximum swarm iterations are shown in Figure \ref{fig:swarmtesting1}.  Other algorithm parameters like $c_1$,$c_2$ that were mentioned in Section\ref{sec:ParticleSwarmAlgorithm} are tested using similar tests (Figure \ref{fig:swarmtesting} and \ref{fig:swarmtesting2}). The final values are for maximum number of iteration = 80 with 15 particles. The final number of agents and maximum swarm iteration used in the system was based trade off between errors achieved versus the computation time.

@@

%The fig. \ref{fig:dataerrorvsiteration.jpg} display the effect of the size of swarm population on the number of vertex reported while segmenting the stroke.\footnote{The fewer the vertex the better the segmentation} The graphs shows that as the number of iteration increase there is a decrease in error until the error reach a saturation and any increase in number of iteration does not affect the error calculated. Similar behavior is noticed in fig. \ref{fig:datapopulation.jpg}.  These curves and test result in choosing the swarm parameters $c_1,c_2$ and maximum number of iteration and the swarm population. The final values are for maximum number of iteration = 150 with 15 particle for better compensation in the time domain.

@@

\subsection{Segmentation Algorithms}
\label{sec:SegmentationAlgorithms}
The result of the segmentation algorithm can be viewed in the Figure \ref{fig:results1}. This result shows the originals stroke and the segmentation that the system generated for the stroke. Figure \ref{fig:results2} shows the result of \textsl{AlgS1} and \textsl{AlgS2} for the same input strokes.  Figure \ref{fig:results3} shows the output of \textsl{Alg3} to the same set of input strokes. It clearly shows that \textsl{AlgS2} have better segmentation results than both \textsl{AlgS1} and \textsl{Alg3} but since these segmentation are subjective to what the user intended to draw the next section will focus on the effect of the segmentation algorithm on the recognition accuracy.

@@

We performed several experiments to evaluate the presented recognition system. Firstly we tested recognition accuracy of shapes in the data set with both algorithms (\textsl{AlgS1} and \textsl{AlgS2}). We also implemented the segmentation algorithm described in \cite{earlyprocess} (\textsl{Alg3}) to use as reference to compare it with our swarm algorithms. Figure \ref{fig:test1} shows the accuracy achieved by each algorithm. The two swarm algorithms were tested with and without the ellipse fitting module. The ellipse detection module appears to be superior. The result shows that combining both  (\textsl{AlgS1} and \textsl{AlgS2}) out perform any single algorithm. %show that combining algoirthmsachieve higher accuracy than any single algoirthm. 
%improves the results.% with both the \textit{DPSO} algorithms.% The results show that both \textit{DPSO} algorithms achieve better result than other algorithms.

@@

The result illustrated in this chapter shows the work done on the system and experiments performed. The results show that the ellipse detection and PSO swarm algorithms gives better performance than other known systems.
